---
editor_options:
  chunk_output_type: console
---

# NOTES

-   This file is *not* included in the [source package](https://r-pkgs.org/structure.html#sec-source-package) because of the [`.nopurl` suffix in its
    filename](https://pkgpurl.rpkg.dev/reference/purl_rmd.html#-rmd-files-excluded-from-purling).

-   The chunks below have to be manually executed in order to regenerate the package data.

-   Although the datasets below are saved as ["internal data" in `R/sysdata.rda`](https://r-pkgs.org/data.html#sec-data-sysdata), they can still be exported and
    documented (by documenting the dataset's quoted name in the main `.Rmd` source file – which only works when the dataset is also `@export`ed), something [not
    explicitly mentioned](https://coolbutuseless.github.io/2018/12/10/r-packages-internal-and-external-data/) in the book [R
    Packages](https://r-pkgs.org/data.html#sec-data-data). To do so, you first need to manually add the `export()` directive in the NAMESPACE file since
    roxygen2 won't add it automatically.

# Define data

## `chunk_opts`

```{r}
chunk_opts <-
  c("src/resources/schema/cell-pagelayout.yml",
    "src/resources/schema/cell-layout.yml") |>
  purrr::map(\(path) {
    
    yay::gh_text_file(path = path,
                      owner = "quarto-dev",
                      name = "quarto-cli") |>
      yaml::read_yaml(text = _) %>%
      magrittr::set_names(value = purrr::map_chr(., \(x) x$name)) |>
      purrr::map(\(x) purrr::pluck(x, "schema", "enum")) |>
      purrr::compact()
  }) |>
  purrr::list_flatten()
```

## `qmd_layout_classes`

```{r}
qmd_layout_classes <-
  yay::gh_text_file(path = "src/resources/schema/definitions.yml",
                    owner = "quarto-dev",
                    name = "quarto-cli") |>
  yaml::read_yaml(text = _) |>
  purrr::keep(\(x) x$id == "page-column") |>
  purrr::chuck(1L, "enum")
```

# Write data

Save all the small data objects as a single internal file `R/sysdata.rda`. Note that when documenting them, they must be explicitly `@export`ed to be available
to package users.

```{r}
usethis::use_data(chunk_opts,
                  qmd_layout_classes,
                  internal = TRUE,
                  overwrite = TRUE,
                  compress = "xz",
                  version = 3L)
```
